---
title: "DeepScanScape Report"
output: html_document
params:
  dms: null
  highlight: null
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.align = "center")
```

```{r load_study, include=FALSE}
dms <- params$dms

if (!dms$imputed) {
  dms <- impute(dms)
}

if (!dms$annotated) {
  dms <- annotate(dms)
}

checks <- check_data(dms, warn = FALSE)
```

This report contains a summary analysis of the following deep mutational scan(s):

`r knitr::kable(dms$meta)`

`r if(nrow(checks) > 0) "# Consistency Checks"`
`r if(nrow(checks) > 0) "These checks identify common abnormalities in deep scan datasets. The following were found:"`
`r if(nrow(checks) > 0) knitr::kable(checks)`

# Fitness Score Distribution

The distribution of fitness scores can tell you properties of the protein(s) studied or indicate incorrectly transformed data.
ER scores on the standard log<sub>2</sub>(ER) generally form a bimodal distribution with a peak of neutral variants near to 0 and a peak of negative deleterious variants.
The overall distribution from the combined landscape dataset is also shown, which can be used to identify deviations.


```{r er_distribution}
plot_er_distribution(dms)
```

# Fitness Heatmap

Viewing the fitness scores across protein positions can identify regions of interest, for example those that are very sensitive to mutations. 

```{r er_heatmap}
plot_er_heatmap(dms)
```

# Mutational Landscape

Mapping new positions onto the deep mutational landscape can also highlight potential properties of the scanned positions.
Here we colour the landscape by most common subtype group, which identifies the general properties of positions that tend to fall in the region.
The `plot_landscape` function can be used to highlight other properties of the landscape as well, for example regions mean SIFT4G score or FoldX $\Delta\Delta G$.

```{r mutational_landscape}
plot_landscape(dms, feature = "cluster", highlight = params$highlight)
```

# Subtypes

The deep mutational landscape can also be divided positional amino acid subtypes, which groups positions of each amino acid into clusters with similar mutational profiles.
New positions can be mapped to these clusters, based on the nearest neighbour cluster centre in PCA space using cosine distance as in the clustering procedure.

```{r subtypes}
plot_cluster_frequencies(dms)
```

These subtypes can inform about the potential properties of positions, both in broad qualitative terms and by making approximate predictions for quantitative properties such as structural destabilisation.
Basic annotations for `r if (is.null(highlight)) "a subset of positons" else "positions of interest"` are shown here, with the full annotation for all positions accessible using the `describe_clusters` function.

```{r subtypes_desc, echo=FALSE}
cols <- c("name", "position", "wt", "cluster", "global_cluster_freq", "group", "description", "notes", "mean_score",
          "total_energy")
desc <- describe_clusters(dms, full = TRUE)

if (is.null(params$highlight)) {
  desc <- desc[1:10, ]
} else {
  highlight <- tibble::tibble(name = rep(names(params$highlight), times = unname(sapply(params$highlight, length))),
                              position = unlist(params$highlight))
  desc <- dplyr::left_join(highlight, desc, by = c("name", "position"))
}

knitr::kable(desc[cols])
```
